---
title: "P8105_hw3_xy2404"
author: "Annie Yu"
date: "10/14/2018"
output: github_document 
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```


# Problem 1:

```{r}
library(p8105.datasets)
data(brfss_smart2010)
brfss =
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor")))
```

The original dataset contains `r colnames(brfss_smart2010)` variables. In original data, there are 23 variables. After data cleaning and focus on the topic of "Overall Health", there are 23 varibales for now dataset. Now the dataset is stored in the dataframe, which has `r nrow(brfss)` rows and `r ncol(brfss)` columns. The new dataset also order the respose from "excellent" to "poor".

## In 2002, which states were observed at 7 locations?
```{r}
brfss %>%
  filter(year == '2002') %>% 
  group_by(locationabbr,locationdesc) %>% 
  summarize(n=n()) %>% 
  group_by(locationabbr) %>% 
  summarize(n=n()) %>% 
  filter(n == 7)

```

In this dataset, there are "CT", "FL", and "NC" which are observed at 7 locations.

## Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
```{r}
brfss %>% 
  group_by(locationabbr, year) %>% 
  summarise(n_obs = length(data_value)) %>% 
  ggplot(aes(x = year, y = n_obs, color = locationabbr))+
  geom_line()+
  ggtitle('The number of locations in each state from 2002 to 2010')+
  theme(legend.position = "bottom")
```

From the spaghetti plot, we could observe the location of FL" is the most.

## Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r}
  ny_excellent = brfss %>% 
  group_by(locationabbr) %>% 
  filter(locationabbr == "NY" & (year == "2002"|year == "2006"|year == "2010")) %>% 
  select(-c(topic, class, question, sample_size, confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 
  group_by(year) %>% 
  summarize(mean_excellent = mean(Excellent),
            sd_excellent = sd(Excellent))

knitr::kable(ny_excellent)
  
```

For the years 2002, the mean of "excellent" responses across locations in NY State is 24.04.
For the years 2006, the mean of "excellent" responses across locations in NY State is 22.53.
For the years 2010, the mean of "excellent" responses across locations in NY State is 22.79.
For the years 2002, the standard deviation of "excellent" responses across locations in NY State is 4.486.
For the years 2006, the standard deviation of "excellent" responses across locations in NY State is 4.001.
For the years 2010, the standard deviation of "excellent" responses across locations in NY State is 3.567.

## Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r}
five_panel <- brfss %>% 
 group_by(year, locationabbr,response) %>% 
  select(-c(topic, class, question, sample_size, confidence_limit_low:confidence_limit_high)) %>% 
  summarise(average_response = mean(data_value, na.rm = TRUE)) 
  
ggplot(five_panel, aes(x = year, y = average_response))+
    geom_point(alpha = 0.5)+
    facet_grid(~response)+
    stat_summary(fun.y = median, geom = "point", color = "blue", size =1)+
    viridis::scale_fill_viridis(discrete = TRUE)+
  labs(
    title = "Distribution of these state-level averages over time",
    x = "Year",
    y = "state level average of proportion"
  )
``` 

The five-panel plot show the highest of average over time is Very good, the lowerest of average over time is Poor. The average over time of Excellent and Good are similar.

# Problem 2

```{r}
library(p8105.datasets)
data(instacart)
```

The dataset has has `r nrow(instacart)` rows and `r ncol(instacart)` columns. The dataset contains `r colnames(instacart)` variables. In the dataset of instacart, the variable of add_to_cart_order means order in which each product was added to the cart. The variable of order_number means the order sequence number of this users. The varibale of aisle means the name of the aisle.I choose the order identifier 1, which the product identifier is 49302 and customer identifier is 112108, and the name of product is Bulgarian Yogurt as one example. At this example, the product is the first order to add the cart and has been ordered by this users in the past. The order of this product is placed 4 times one week and it is 10 times the hour of the day, The aisle of this product is in yogurt and department is dairy eggs.

## How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% 
  distinct(aisle) %>% 
  nrow()

tail(names(sort(table(instacart$aisle))),1)
```

There are 134 aisles in this dataset, and the most items are orderd from the "fresh vegetables".


## Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>% 
  group_by(aisle_id) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = aisle_id, y = n)) + 
  geom_point() + 
  labs(
    title = "Aisle distribution plot",
    x= "Aisle_ID",
    y= "Number of orders placed"
  )
```

I make the scatteplot for this question, the x-axis is Aisle_ID and the y-axis is the number of order placed. This plot has shown there are two highest point. 

## Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
```{r}
most_ais = instacart %>% 
  group_by(aisle, product_id, product_name) %>% 
  filter(aisle == "baking ingredients"|aisle == "dog food care"|aisle == "packaged vegetables fruits") %>% 
  count() %>% 
  group_by(aisle) %>% 
  arrange(n) %>% 
  filter(min_rank(desc(n)) ==1)
knitr::kable(most_ais)
```

The most popular items aisles "god food care" has 30 and its product name is "Snack Sticks Chincken and Rice precipe Dog Treats".
The most popular items aisles "baking ingredients" has 499 and its product name is "Light Brown Suger".
The most popular items aisles "Package vegetables firuts" has 9784 and its product name is "Organic Baby Spinach". And it is the most aisles.


## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
```{r}
mean_instacart = instacart %>% 
  select(order_hour_of_day, product_name, order_dow) %>% 
  filter(product_name == "Pink Lady Apples"| product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour)

 knitr::kable(mean_instacart) 
```

From ths table, we can conclude the order of the mean hour of the day at "Coffee Ice Cream" is more than the mean hour of the day at "Pink Lady Apples".


# Problem 3
```{r}
library(p8105.datasets)
data(ny_noaa)
```

The dataset is from the NOAA National Climatic Data Center. it showed the maximum and minimum temperature, total daily precipitation, snowfall and snow depth. The dataset has has `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The dataset contains `r colnames(ny_noaa)` variables. Because there are missing data(NA), we need to remove NA first if we hope to do some data processing.

## Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
nynoaa <- 
  ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into = c("Year", "Month", "Day"), sep = "-") %>% 
  janitor::clean_names() 

knitr::kable(nynoaa %>%   
  group_by(snow) %>% 
  summarise(n=n()) %>% 
  mutate(tem_ranking =  min_rank(desc(n))) %>% 
  filter(min_rank(desc(n)) ==1))
  
```

The most commonly observed values is 0, because most of time is no snowfall. 

## Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
ny_df_1 =
  nynoaa %>% 
  group_by(month,id,year) %>% 
  mutate(tmax = as.numeric(tmax)) %>% 
  mutate(tmin = as.numeric(tmin)) %>% 
  filter(month == "01"|month == "07") %>%
  na.omit() %>% 
  summarise(mean_temp = mean(tmax*0.1))

```

```{r}
ny_df_1 %>% 
ggplot(aes(y = mean_temp))+
    geom_boxplot()+
    facet_grid(.~month)+
    ggtitle("The average max temperature in January and in July in each station across years")
```

```{r}
ny_df_1 %>% 
ggplot(aes(x = year, y = mean_temp))+
   geom_point()+
   ggtitle("The average max temperature in January and in July in each station across years")+
   facet_grid(.~month)+
   stat_summary(fun.y = median, geom = "point", color = "blue", size =1)+
   viridis::scale_fill_viridis(discrete = TRUE)
```

In January, there are outliers and In July, there are also outliers. From the boxplot, we can conclude the average maximum temperature in July is higher than thenaverage maximum temperature in January becasue it is summer in July.  


## Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option).(2)Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r}
tmax_vs_tmin <-nynoaa %>% 
  group_by(year, month) %>% 
  na.omit(tmax) %>% 
  na.omit(tmin) %>% 
  mutate(tmax_2 = as.numeric(tmax, na.rm = TRUE)) %>% 
  mutate(tmin_2 = as.numeric(tmin, na.rm = TRUE)) %>% 
  ggplot(aes(tmin_2,tmax_2))+
  geom_hex(bins = 15)+
  theme(legend.position = "bottom")

snowfall_value<-nynoaa %>% 
  filter(snow>0, snow<100) %>% 
  ggplot(aes(x = snow, fill = year))+
  geom_density(alpha = 0.5)+
  ggtitle("The distribution of snowfall")+
  theme(legend.position = "bottom")

  tmax_vs_tmin+snowfall_value  

```

The plot of tmax vs tmin show a linear relationship. When the tmin is icreasing, the tmax is increaing too. According to commpare the temperature plot and the distribution of snowfall values, we can conclude when the temperature is lowest, the snowfall values is the most. When the temperature is highest, there is no snowfall. 